---
title: "TUM-RC coefficients"
author: "Oksana Chernova"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    toc: true
    toc_depth: 2
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This RMarkdown file contains the TUM-RC model coefficients used to predict clinically significant prostate cancer (defined as ISUP ≥ 2).

#  Loading libraries
```{r , message = FALSE, warning=FALSE}
library(knitr)
library(dplyr) 
library(tidyr)
library(stringr)
```

```{r  echo=FALSE, results = 'hide'}
library(fmtr)
# Create formatting list
formatnames <- flist(char = value(condition(x == "age", "Age"),
                                  condition(x == "(Intercept)", "Intercept"),
                                  condition(x == "PIRADS", "PIRADS"),
                                  condition(x == "PIRADScat24", "PIRADS 4"),
                                  condition(x == "PIRADScat25", "PIRADS 5"),
                                  condition(x == "PIRADScat34-5", "PIRADS 4-5"),
                                  condition(x == "log2_psa", "log2 PSA"),
                                  condition(x == "log2_vol", "log2 Volume"),
                                  condition(x == "DREabnormal", "DRE abnormal"),
                                  condition(x == "Prior_neg_biopsyYes", "Prior negative biopsy Yes"),
                                  condition(x == "priornegbiopsyyes", "Prior negative biopsy Yes")
))

sumreg <- function(rg){
  t <- summary(rg)
  estim <- round(t$coef[, 1], 4) 
  HR <- format(round(exp(t$coef[, 1]), digits=2), 2, nsmall=2) 
  HR.confint.lower <- round(exp(t$coef[, 1] + qnorm(0.025) * t$coef[,2]), digits=2)
  HR.confint.upper <- round(exp(t$coef[, 1] + qnorm(0.975) * t$coef[,2]), digits=2)
  
  HR <- paste0(HR, " (", HR.confint.lower, ", ", HR.confint.upper, ")")
  p.value <- signif(t$coef[, 4], digits=4)
  
  t2 <- data.frame(estim, HR, p.value)#estim
  t2$p.value <- ifelse(t2$p.value < 0.001, '< 0.001',format(round(t2$p.value, 3), nsmall = 3))
  colnames(t2) <- c('Estimate', 'OR (95% CI)', 'P value')
  # apply format
  rownames(t2) <- fapply(rownames(t2), formatnames)
  
  return(kable(t2, format = 'markdown'))
}
```

# Read data
```{r}
data <- read.csv("C:/Users/ochernova/mydocs/pr2/MRIbiopsy21.01.2026.csv")  

dim(data) 

# Define clinically significant prostate cancer as ISUP >= 2
data <- data %>% 
  mutate(
    PSA = as.numeric(psa),
    PIRADS = as.numeric(PIRADS),
    csPCa = as.factor(ifelse(ISUP >= 2, "csPCa", "non-csPCa")),
    csPCa2 = ifelse(ISUP >= 2, 1, 0),
    log2_psa = log2(psa),
    log2_vol = log2(volume)
  )

# Define factor levels ----
data <- data %>% mutate(
  priornegbiopsy = relevel(factor(priornegbiopsy), ref="no")
  ,DRE = relevel(factor(DRE), ref="normal")
)

data = data[,c("age", "log2_psa", "log2_vol", "PIRADS", "DRE", "priornegbiopsy", "csPCa2")]
```

# Missing values
```{r}
na_counts <- colSums(is.na(data))
print(na_counts)
```

# Model for complete cases
```{r}
data_complete <- data %>%
  filter(!is.na(DRE) & !is.na(log2_vol) & !is.na(PIRADS) & !is.na(priornegbiopsy))

print(paste0("sample size: ", nrow(data_complete)))
      
model1 <- glm(formula = csPCa2 ~  age + log2_psa + PIRADS + log2_vol + DRE + priornegbiopsy,
              data = data_complete,
              family=binomial(link='logit'))

sumreg(model1)

paste(as.character(round(summary(model1)$coef[, 1], 5)), sep="' '", collapse=", ")
```

# Model without priornegbiopsy
```{r}
data_bx <- data %>%
  filter(!is.na(DRE) & !is.na(log2_vol) & !is.na(PIRADS) )

print(paste0("sample size: ", nrow(data_bx)))
      

model2 <- glm(formula = csPCa2 ~  age + log2_psa + PIRADS  + log2_vol + DRE ,
              data = data_complete,
              family=binomial(link='logit'))

sumreg(model2)
paste(as.character(round(summary(model2)$coef[, 1], 5)), sep="' '", collapse=", ")
```

# Model without DRE
```{r}
data_dre <- data %>%
  filter(!is.na(log2_vol) & !is.na(priornegbiopsy) & !is.na(PIRADS))

print(paste0("sample size: ", nrow(data_dre)))

model3 <- glm(formula = csPCa2 ~  age + log2_psa + PIRADS + log2_vol + priornegbiopsy,
              data = data_dre,
              family=binomial(link='logit'))

sumreg(model3)
paste(as.character(round(summary(model3)$coef[, 1], 5)), sep="' '", collapse=", ")
```


# Model without volume
```{r}
data_volume <- data %>%
  filter(!is.na(DRE) & !is.na(priornegbiopsy) & !is.na(PIRADS))

print(paste0("sample size: ", nrow(data_volume)))

model4 <- glm(formula = csPCa2 ~  age + log2_psa + PIRADS + DRE + priornegbiopsy ,
              data = data_volume,
              family=binomial(link='logit'))

sumreg(model4)
paste(as.character(round(summary(model4)$coef[, 1], 5)), sep="' '", collapse=", ")
```


# Model without priornegbiopsy and DRE
```{r}
data_bxdre <- data %>%
  filter( !is.na(log2_vol) & !is.na(PIRADS) )

print(paste0("sample size: ", nrow(data_bxdre)))

model5 <- glm(formula = csPCa2 ~  age + log2_psa + PIRADS + log2_vol,
              data = data_bxdre,
              family=binomial(link='logit'))

sumreg(model5)
paste(as.character(round(summary(model5)$coef[, 1], 5)), sep="' '", collapse=", ")
```


# Model without priornegbiopsy and volume
```{r}
data_bxvol <- data %>%
  filter(!is.na(DRE) & !is.na(PIRADS) )

print(paste0("sample size: ", nrow(data_bxvol)))

model6 <- glm(formula = csPCa2 ~  age + log2_psa + PIRADS + DRE,
              data = data_bxvol,
              family=binomial(link='logit'))

sumreg(model6)
paste(as.character(round(summary(model6)$coef[, 1], 5)), sep="' '", collapse=", ")
```

# Model without volume and DRE
```{r}
data_volumedre <- data %>%
  filter(!is.na(priornegbiopsy) & !is.na(PIRADS))

print(paste0("sample size: ", nrow(data_volumedre)))

model7 <- glm(formula = csPCa2 ~  age + log2_psa + PIRADS + priornegbiopsy,
              data = data_volumedre,
              family=binomial(link='logit'))

sumreg(model7)
paste(as.character(round(summary(model7)$coef[, 1], 5)), sep="' '", collapse=", ")
```

# Model without DRE, volume, priornegbiopsy
```{r}
data_volumedrebx <- data %>%
  filter(!is.na(PIRADS))

print(paste0("sample size: ", nrow(data_volumedrebx)))

model8 <- glm(formula = csPCa2 ~  age + log2_psa + PIRADS,
              data = data_volumedrebx,
              family=binomial(link='logit'))

sumreg(model8)
paste(as.character(round(summary(model8)$coef[, 1], 5)), sep="' '", collapse=", ")
```